---
title: "HW2_Tingyang_Hu"
author: "Tingyang Hu"
date: "9/23/2021"
output: 
  html_document: 
    theme: sandstone
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

The dataset contains the Covid-19 vaccination data across countries, territories and areas provided by WHO to the data of 08/31/2021. The variables include countries, cumulative total vaccine doses administered, cumulative number of persons vaccinated with at least one dose, vaccine types used, etc. Exploratory analysis including characterizing summary statistics and data visualization will be performed on the data. 

## Loading and preparing data
```{r data, echo = FALSE}
# read in data
vaccine <- read.csv("~/Desktop/vaccination-data.csv",header = T)

knitr::kable(head(vaccine[,-12]),
             caption = "Table 1:The first 6 rows of the vaccine dataset")

# look at the descriptive statistics 
knitr::kable(summary(vaccine),
             caption = "Tabl 2:The Summary Statistics the vaccine dataset")

# deal with missing values: impute with mean of variables
fun1<- function(x) {x[is.na(x)] <- mean(x, na.rm = TRUE)
                    x}
vaccine$TOTAL_VACCINATIONS_PER100 <- fun1(vaccine$TOTAL_VACCINATIONS_PER100)
vaccine$PERSONS_FULLY_VACCINATED_PER100 <- fun1(vaccine$PERSONS_FULLY_VACCINATED_PER100)
```

## Figures 

The figure for vaccination across different countries are showed, in which the shade of the color indicates number of vaccinations.
```{r load packages, include=FALSE}
library (maps)
library (mapdata)
library(ggplot2)
library(dplyr)
```


```{r figure1, echo=FALSE, fig.cap="Figure 1:Vaccinated Population Distribution"}
# map with distribution of vaccination

world_map<-map_data("world")
 
#deal with country variables
vaccine$COUNTRY<- gsub("United States of America","USA",vaccine$COUNTRY)
vaccine$COUNTRY<- gsub("The United Kingdom","United Kingdom",vaccine$COUNTRY)
vaccine$COUNTRY<- gsub("Russian Federation","Russia",vaccine$COUNTRY)

# merge data for map
world_map %>% 
  left_join(vaccine, by = c("region" = "COUNTRY")) -> vaccine_world_map

#head(vaccine_world_map)

# plot the map with total vaccination 
 ggplot(vaccine_world_map, aes(x = long, y = lat, group = group)) +
  geom_polygon(aes(fill= TOTAL_VACCINATIONS_PER100), colour = "white") +
  scale_x_continuous(breaks = seq(-180, 210, 45), labels = function(x){paste0(x, "°")}) +
  scale_y_continuous(breaks = seq(-60, 100, 30), labels = function(x){paste0(x, "°")}) +
  scale_fill_gradient(low = "lightblue", high="steel blue") +
  labs(title="Total Vaccinations per 100 Across Countries", y="Latitude", x="Longitude") + theme_light() +
  theme(legend.key.size = unit(18, "pt"),legend.title=element_text(size=8),legend.position = "bottom") + coord_fixed() 

```
## My working path and info

The working directory and session information in my laptop are as follows:
```{r wd,echo=FALSE}
getwd() 
devtools::session_info()
```

